The LHC as a Neutrino-Ion Collider

### 2023-11-27 referee/referee-reply.tex

\end{enumerate}

\noindent
We now address the comments of referee B

\begin{enumerate}
\item{\it Page 3, lines 34-36 column 2 “since in general the strange and charm … not expected to vanish”. I agree but I suggest adding a reference or a comment supporting this statement. }

  We have added the following references which explore strange and charm PDF asymmetries ~\cite{Sufian:2018cpj, Sufian:2020coz}.
  %
  We have also added a reference to the recent NNPDF study of the intrinsic charm asymmetry in the proton,
  showing that these effects are potentially sizable and could hence be relevant for the interpretation
  of the FPF measurements.

@@

\item{\it p. 4 l. 30-35 col.2 “Also, to identify… event yields.” Why a cut on the hadronic energy should properly simulate a cut on the number of charged tracks emerging from the interaction vertex?}

The charged track multiplicity is expected to grow with the hadronic system’s invariant mass. We make this clear and cite a neutrino-Hydrogen interaction multiplicity study ~\cite{Aachen-Bonn-CERN-Munich-Oxford:1981lfk} as well as a FASER paper ~\cite{FASER:2019dxq} which also point this out. We also append this sentence:

### 2023-11-27 referee/referee-reply.tex

	Following the referee's suggestion, the results which only include the statistical errors
	are now labelled "FPF$\star$" while the one that also account for the systematics are labelled "FPF".
        %
        We have also extended the rationale for showing results in the statistics-only scenario, alongside the
        considerations described above.

	\item {\it Section 4 and elsewhere. Given these are HL-LHC projections, somewhere these should
		be compared with the HL-LHC PDFs of Ref [34]. This would surely be the fairer
		comparison, or in any case will give a clearer picture of where things may stand.
	}

	The reason why we did not provide comparisons with the HL-LHC PDFs of Ref. [34] is because
	the pseudodata used to determine these PDFs were produced with the PDF4LHC15 set which
	does not account for the recent LHC measurements. The most consistent comparisons would
	be to re-do the analysis done in ref. [34] using PDF4LHC21 and then compare the resulting PDFs 
	with our determination, something which is clearly outside the scope of this paper.

        This said, to answer the question from the referee, we have repeated the phenomenological study of Sect 4
        comparing PDF4LHC15 with PDF4LHC15+HLHC PDFs for the same processes.
        %
        Specifically, we have added to the paper an Appendix where the phenomenology study of Sect 4 is repeated
        now for the PDF4LHC15 with PDF4LHC15+HLHC PDFs.
        %
        From this comparison, one can observe a significant reduction in the PDF uncertainties, especially for
        processes that involve single gauge boson productions. The decrease in PDF uncertainties, for most of
        the processes and most of the kinematic regions, are greater than the decrease observed in FPF with only
        statistical uncertianties.

### 2023-11-27 journal_sub/app-faser.tex
Add epjc typeset version
Add HL-LHC phenos and move FASERv results to Sec. 3

\section{Impact of FASER$\nu$ Run III measurements}
\label{app:fasernu_runIII_impact}

The same procedure applied to the FPF experiments in Sect.~\ref{sec:pdf4lhc21}
can also be used in the case of FASER$\nu$, for which pseudo-data is generated
for a Run III integrated
luminosity of $\mathcal{L}=150$ fb$^{-1}$.
%
As indicated by Table~\ref{tab:integrated_rates}, by the end of Run III, one expects
that FASER$\nu$ will have recorded around 600 and 1800 electron- and muon-neutrinos
respectively corresponding to deep-inelastic scattering events, of which around 100
and 300 respectively are associated to charm-tagged events.
%
These projected event rates can be compared with the expectations for FASER$\nu$2, which predict
an increase in statistics of more than two orders of magnitude.

Fig.~\ref{fig:FASERnu2_vs_FASERnu} presents a similar comparison as that of
Fig.~\ref{fig:FASERnu2_baseline} now between the projected PDF impact of FASER$\nu$2 and that of
FASER$\nu$.
%
One finds that neutrino DIS measurements at FASER$\nu$ are unable to improve
PDF uncertainties as compared to the baseline scenario encapsulated by PDF4LHC21.
%
The reason is two-fold: the smaller event rates as compared to FASER$\nu$2,
and the reduced coverage of the $(x,Q^2)$ phase space shown in Fig.~\ref{fig:fasernu2_muon}.
%
The differences in PDF sensitivity between FASER$\nu$ and FASER$\nu$2 in
Fig.~\ref{fig:FASERnu2_vs_FASERnu} illustrate
the  importance of realising the FPF in order to exploit the full physics potential enabled 
by LHC neutrinos for QCD and hadron structure studies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_dv_ratio.pdf}
\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_g_ratio.pdf}\\
\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_s_ratio.pdf}
\caption{Same as  Fig.~\ref{fig:FASERnu2_baseline} (statistical uncertainties only), comparing
  the projected PDF impact of FASER$\nu$2 with that of FASER$\nu$, in the latter
  case assuming the Run III integrated
luminosity of $\mathcal{L}=150$ fb$^{-1}$.
}
\label{fig:FASERnu2_vs_FASERnu}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

It should be emphasized that the lack of PDF sensitivity displayed in Fig.~\ref{fig:FASERnu2_vs_FASERnu}
does not imply that measuring DIS structure functions at FASER$\nu$ will not provide useful information.
%
First of all, our procedure assumes perfect compatibility between current data and the projected DIS
neutrino measurements at the LHC, something which remains to be demonstrated experimentally.
%
Second, FASER$\nu$ still covers a region of neutrino energies unexplored by previous experiments, hence providing
a new validation of our QCD calculations and of neutrino interactions at the TeV scale.
%
Third, it would represent a non-trivial proof-of-concept that LHC neutrino differential measurements can be
unfolded to the cross-section level to be used in  theoretical
interpretations, paving the way and demonstrating the feasibility of subsequent neutrino DIS
measurements at the FPF experiments.

>>>>

\section{Phenomenology of HL-LHC}
\label{app:comparisons_with_HLLHC}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!h]
	\centering
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_DY_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WP_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HVBF_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HWP_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HWM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WPWM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WPZ_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WMZ_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\caption{Same as}
	\label{fig:HLLHC_pheno_integrated}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the following section, we repeat the same phenomenology studies as in
Sec.~\ref{sec:pheno} but using the High-Luminosity LHC (HL-LHC) PDF 
from~\cite{AbdulKhalek:2018rok}.
%
Such comparisons present a complementary study to the impact seen both at the PDF 
and prediction level shown in Sec.~\ref{sec:pheno}. The PDF constraints from the 
FPF and HL-LHC experiments are fully orthogonal in that the former have the very 
important property that possible contamination from new physics on the PDF fits 
can be completely ignored, since this process is driven by $Q^2$ values outside 
the possible presence of BSM physics.
%
Specially, should anomalies be revealed at the HL-LHC, having the fully independent
validation of the large-$x$ PDFs provided by the FPF would be extremely valuable 
for its interpretation.

In Fig.~\ref{fig:HLLHC_pheno_integrated} are displayed the fiducial cross-sections for 
representative LHC processes, similar to the ones shown in Fig.~\ref{fig:NNPDF40_pheno_integrated},
but evaluated with PDF4LHC15 and its variant that includes HL-LHC projections (``\_HLLHC'').
The corresponding differential cross-sections are shown in Fig.~\ref{fig:HLLHC_pheno_differential}.
%
Concerning the fiducial cross-sections, a significant reduction in PDF uncertainties
is seen, reaching up to a factor of four for processes involving single boson productions. 
Such a significant decrease in the PDF uncertainties is also observed at the differential
level.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htbp]
	\centering
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_DY_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WP_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HVBF_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HWP_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HWM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WPWM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WPZ_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WMZ_14TEV_40_PHENO-global-HLLHC.pdf}
	\caption{Same as Fig.~\ref{fig:NNPDF40_pheno_integrated}
		for the corresponding differential distributions in the case of NNPDF4.0
		%
	}
	\label{fig:HLLHC_pheno_differential}
\end{figure*}

### 2023-11-27 journal_sub/sec-results.tex
Add HL-LHC phenos and move FASERv results to Sec. 3

By following the strategy outlined in Sect.~\ref{sec:dis_pseudodata}, we 
quantify the impact on the proton and nuclear PDFs of differential  DIS
cross-section measurements obtained with the  LHC neutrino beam. 
%
Here we present results first for the Hessian profiling of the PDF4LHC21,
then for the Monte Carlo fit NNPDF4.0, and finally for the nuclear PDFs of EPPS21, also
by means of profiling.
%
We study the stability of the results with respect to the inclusion of systematic uncertainties,
charm-tagged data, and lepton-charge separation.
%
We also compare the impact of the different FPF experiments separately and provide
results for their combination.
%
For completeness, we study in App.~\ref{app:fasernu_runIII_impact} the constraints on PDFs 
provided by the forward experiments, namely FASER$\nu$. The results show that measurements
at FASERV$\nu$ are unable to improve the determination of PDFs therefore strengthening the
need for the FPF experiments.

@@

>>>>

By following the strategy outlined in Sect.~\ref{sec:dis_pseudodata}, we 
quantify the impact on the proton and nuclear PDFs of differential  DIS
cross-section measurements obtained with the  LHC neutrino beam. 
%
Here we present results first for the Hessian profiling of the PDF4LHC21,
then for the Monte Carlo fit NNPDF4.0, and finally for the nuclear PDFs of EPPS21, also
by means of profiling.
%
We study the stability of the results with respect to the inclusion of systematic uncertainties,
charm-tagged data, and lepton-charge separation.
%
We also compare the impact of the different FPF experiments separately and provide
results for their combination.
%
For completeness, we study the constraints on PDFs 
provided by the forward experiments, namely FASER$\nu$. The results show that measurements
at FASER$\nu$ are unable to further improve the determination of PDFs therefore strengthening the
need for the FPF experiments.

@@

\paragraph{Impact of FASER$\nu$ Run III measurements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htbp]
	\centering
	\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_uv_ratio.pdf}
	\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_dv_ratio.pdf}
	\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_g_ratio.pdf}\\
	\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_Sea_ratio.pdf}
	\includegraphics[width=0.32\textwidth]{plots/proton_fasernu2/FASERv2_vs_FASERv/statOnly_FASERv_q2_10000_pdf_s_ratio.pdf}
	\caption{Same as  Fig.~\ref{fig:FASERnu2_baseline} (statistical uncertainties only), comparing
		the projected PDF impact of FASER$\nu$2 with that of FASER$\nu$, in the latter
		case assuming the Run III integrated
		luminosity of $\mathcal{L}=150$ fb$^{-1}$.
	}
	\label{fig:FASERnu2_vs_FASERnu}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The same procedure applied to the FPF experiments in earlier
can also be used in the case of FASER$\nu$, for which pseudo-data is generated
for a Run III integrated
luminosity of $\mathcal{L}=150$ fb$^{-1}$.
%
As indicated by Table~\ref{tab:integrated_rates}, by the end of Run III, one expects
that FASER$\nu$ will have recorded around 600 and 1800 electron- and muon-neutrinos
respectively corresponding to deep-inelastic scattering events, of which around 100
and 300 respectively are associated to charm-tagged events.
%
These projected event rates can be compared with the expectations for FASER$\nu$2, which predict
an increase in statistics of more than two orders of magnitude.

Fig.~\ref{fig:FASERnu2_vs_FASERnu} presents a similar comparison as that of
Fig.~\ref{fig:FASERnu2_baseline} now between the projected PDF impact of FASER$\nu$2 and that of
FASER$\nu$.
%
One finds that neutrino DIS measurements at FASER$\nu$ are unable to improve
PDF uncertainties as compared to the baseline scenario encapsulated by PDF4LHC21.
%
The reason is two-fold: the smaller event rates as compared to FASER$\nu$2,
and the reduced coverage of the $(x,Q^2)$ phase space shown in Fig.~\ref{fig:fasernu2_muon}.
%
The differences in PDF sensitivity between FASER$\nu$ and FASER$\nu$2 in
Fig.~\ref{fig:FASERnu2_vs_FASERnu} illustrate
the  importance of realising the FPF in order to exploit the full physics potential enabled 
by LHC neutrinos for QCD and hadron structure studies.

It should be emphasized that the lack of PDF sensitivity displayed in Fig.~\ref{fig:FASERnu2_vs_FASERnu}
does not imply that measuring DIS structure functions at FASER$\nu$ will not provide useful information.
%
First of all, our procedure assumes perfect compatibility between current data and the projected DIS
neutrino measurements at the LHC, something which remains to be demonstrated experimentally.
%
Second, FASER$\nu$ still covers a region of neutrino energies unexplored by previous experiments, hence providing
a new validation of our QCD calculations and of neutrino interactions at the TeV scale.
%
Third, it would represent a non-trivial proof-of-concept that LHC neutrino differential measurements can be
unfolded to the cross-section level to be used in  theoretical
interpretations, paving the way and demonstrating the feasibility of subsequent neutrino DIS
measurements at the FPF experiments.
